---
title: "Discovering regulated Metabolite Families"
author:
- name: Khabat Vahabi
  affiliation: Leibniz Institute of Plant Biochemistry
- name: Steffen Neumann
  affiliation: Leibniz Institute of Plant Biochemistry
  email: sneumann@ipb-halle.de
package: MetFamily
output:
  BiocStyle::html_document:
    toc_float: true
abstract: |
  Description of the MetFamily R package
vignette: |
  %\VignetteIndexEntry{Discovering regulated Metabolite Families with MetFamily}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
csl: biomed-central.csl
---

```{r setup, include=FALSE, echo=FALSE}
library(MetFamily)
```

# Introduction

Some text about the scenario, data and MTBLS297.

Cool paper [@Treutler16DiscoveringRegulatedMetabolite].
 
Cool R package: `r Githubpkg("ipb-halle/MetFamily")`



# Loading the data

First, we load the data and summarise it

```{r load-data}

fileName <- system.file("extdata/showcase/Project_file_showcase_annotated.csv.gz", package = "MetFamily")

project <- readClusterDataFromProjectFile(file = fileName)

```

# Filtering data

We can filter the data to remove low quality data points

```{r filter-data}

# todo...

```


# PCA

PCA on MS1 in Figure \@ref(fig:pca). 

```{r pca, fig.cap="PCA of MS1.", echo=FALSE}
plot(cars)
```

# HCA

HCA on MS2 in Figure \@ref(fig:hca). 

```{r hca, fig.cap="HCA on MS2.", echo=FALSE}

# Something with calcPlotDendrogram_plotly() 

plot(cars)
```

# References {.unnumbered}

<div id="refs"></div>

# (APPENDIX) Appendix {-} 

## Session info

```{r sessionInfo, echo=FALSE}
sessionInfo()
```

